### 2015-03-11 doc/Poster/example.tex
modified analysis. Made poster

% An example document showing the use of the ucbposter, based on the
% infposter demo developed by M.Reddy (MXR), 24/2/97 and updated by
% Mary E Foster, 10/5/2002.

\documentclass{ucbposter}

\let\AuthorColour\red

%http://www.inf.ed.ac.uk/systems/tex/infposter/infposter.html
%\epsfigure[<width-percent>]{<filename>}{<caption>}

\begin{document}

\title{An Example of the UCB Poster Class}
%\subtitle{A simple title for a simple poster.}
\author{Your Name(s)}
\email{an@address}
%\homepage{http://radlab.cs.berkeley.edu/}
\makeposter

\section{Poster content}

\begin{itemize}
  \item An example bullet point
  \begin{itemize}
    \item Example
  \end{itemize}
\end{itemize}

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.

\section{Poster content}

An example showing the use of the ucbposter style. This style is
heavily based on the University of Edinburgh infposter style. The
original version of this demo were developed by M.Reddy (MXR), 24/2/97
and updated by Mary E Foster, 10/5/2002. Please find more information
regarding the original style at {\tt
http://www.inf.ed.ac.uk/systems/tex/infposter/}.
\end{document}